---
title: Analysis of space audit data
author: Gibran Hemani
date: "`r Sys.Date()`"
output: html_document
---

# Analysis of space audit data

For three weeks from early June a space audit was conducted at 1-5WLR, OH and CH. The purpose of the space audit was to record which areas were being used, how they were being used and when.

Data was collected with

- door sheets, aimed to collect how the office was being used
- walkthroughs, where someone counted how many people were in each office space 3 times a day
- Planon records

The main objective of this analysis is to estimate how many offices will be required at Augustine's Court based on current usage.

There are some limitations with the data.

Door sheets
- Quite incomplete. If we analyse all door data it will be an underestimate of office usage. If we restrict to only offices with data it will be an overestimate.
- Incomplete data about types of meetings
- Some data didn't record dates etc.

Walkthroughs
- Didn't record type of activities in the offices especially if meeting rooms were used for video calls.
- Is a sample which may not be representative of the whole day.

Planon records
- Many people don't use these data
- Bookings were often not confirmed, which may indicate that the office was not used or could just be because it's too onerous to confirm a booking.


## Analysis

### Exploration

Load libraries

```{r}
library(dplyr)
library(readxl)
library(ggplot2)
library(lubridate)
library(tidyr)
library(here)
library(purrr)
```

Read in the Oakfield House walkthrough data
- Convert to long format
- Remove "NOT IN USE" values
- Extract date and time from the name column
- Extract max number of people for each room

```{r}
expand.grid(floor=c("GF","1st", "2nd"), week=c(1,2)) %>% mutate(sheet_name = paste0(floor, " Week ", week)) %>%
    purrr::pmap(., function(floor, week, sheet_name) {
      read_xlsx(here("data/Oakfield House June 2025 for sharing.xlsx"), sheet = sheet_name, skip=1) %>%
        pivot_longer(cols=-c("ROOM NUMBERS", "USE")) %>%
        filter(value != "NOT IN USE", !is.na(`ROOM NUMBERS`)) %>%
        mutate(
            max_number=as.numeric(gsub("[^0-9]", "", USE)), 
            week=week, 
            floor=floor, 
            value=as.numeric(value),
            name=gsub("3.00", "15.00", name) %>% gsub("09.30", "10.00", .) %>% gsub("\\.", ":", .),
            datetime=as.POSIXct(name, format="%d/%m/%Y %H:%M"),
            date=as.Date(datetime, format="%d/%m/%Y"),
            time=format(datetime, "%H:%M"),
            day=wday(datetime, label=TRUE)
        ) %>%
        filter(!`ROOM NUMBERS` %in% c("TOTALS", "multi occupancy space", "Single occupancy office", "Meeting space", ""), !grepl("Meeting", `ROOM NUMBERS`))
    }) %>% bind_rows -> oh_wt
oh_wt %>% str
```

```{r}
table(is.na(oh_wt$max_number))
table(is.na(oh_wt$value))
```

Make a heatmap of the occupancy of Oakfield House by floor and time

```{r}
oh_wt %>% filter(!is.na(max_number)) %>%
    mutate(
        occupancy_ratio=pmin(value/max_number, 1),
        office_type=ifelse(max_number == 1, "Single", "Multi")
    ) %>%
ggplot(., aes(y=as.factor(datetime), x=`ROOM NUMBERS`, fill=occupancy_ratio)) +
  geom_tile() +
  facet_grid(. ~ office_type, scale="free_x", space="free_x") +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(title="Occupancy of Oakfield House by floor and time",
       y="Date and Time",
       x="Room Number",
       fill="Occupancy Ratio") +
       theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

By date

```{r}
oh_wt %>% group_by(date, `ROOM NUMBERS`, floor) %>%
    summarise(
        max_number=mean(max_number, na.rm=TRUE),
        value=max(value, na.rm=TRUE),
        occupancy_ratio=pmin(value/max_number, 1),
        office_type=ifelse(max_number == 1, "Single", "Multi")
    ) %>%
    filter(!is.na(occupancy_ratio)) %>%
    ggplot(., aes(y=as.factor(date), x=`ROOM NUMBERS`, fill=occupancy_ratio)) +
        geom_tile() +
        facet_grid(. ~ office_type, scale="free_x", space="free_x") +
        scale_fill_gradient(low = "white", high = "blue") +
        labs(title="Occupancy of Oakfield House by floor and time",
            y="Date and Time",
            x="Room Number",
            fill="Occupancy Ratio") +
       theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

By day of week

```{r}
oh_wt %>% group_by(day, `ROOM NUMBERS`, floor) %>%
    summarise(
        max_number=mean(max_number, na.rm=TRUE),
        value=mean(value, na.rm=TRUE),
        occupancy_ratio=pmin(value/max_number, 1),
        office_type=ifelse(max_number == 1, "Single", "Multi")
    ) %>%
    filter(!is.na(occupancy_ratio)) %>%
    ggplot(., aes(y=day, x=`ROOM NUMBERS`, fill=occupancy_ratio)) +
        geom_tile() +
        facet_grid(. ~ office_type, scale="free_x", space="free_x") +
        scale_fill_gradient(low = "white", high = "blue") +
        labs(title="Occupancy of Oakfield House by floor and time",
            y="Date and Time",
            x="Room Number",
            fill="Occupancy Ratio") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
oh_wt %>% 
    group_by(time, `ROOM NUMBERS`, floor) %>%
    summarise(
        max_number=mean(max_number, na.rm=TRUE),
        value=mean(value, na.rm=TRUE),
        occupancy_ratio=pmin(value/max_number, 1),
        office_type=ifelse(max_number == 1, "Single", "Multi")
    ) %>%
    filter(!is.na(occupancy_ratio)) %>%
    ggplot(., aes(y=time, x=`ROOM NUMBERS`, fill=occupancy_ratio)) +
        geom_tile() +
        facet_grid(. ~ office_type, scale="free_x", space="free_x") +
        scale_fill_gradient(low = "white", high = "blue") +
        labs(title="Occupancy of Oakfield House by floor and time",
            y="Date and Time",
            x="Room Number",
            fill="Occupancy Ratio") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


How many meeting rooms do we need?

To work this out

- For any date/time, find out how many single ocupancy offices or multi-occupancy meeting rooms are occupied
- Need a list of rooms to be used
- Ignore the staff room, seminar room

```{r}
meeting_romms <- c("OG18","OG19a","OF1A","OF1B","OF2","OF3","OF4","OF5","OF6","OF7","OF8","OF9","OF10.1","OF10.2","OF18","OF19","OF20","OF21","OF22","OF23","OF24","OF26","OF27","OS2A","OS2B","OS9","OS10","OS11","OS12","OS13","OS14","OS15","OS16","OS17.1","OS17.2","OS17.3")

of_wt_mr <- subset(oh_wt, `ROOM NUMBERS` %in% meeting_romms)

of_wt_mr %>%
    mutate(
        in_use=as.numeric(value > 0)
    ) %>%
    group_by(datetime, day, time, date) %>%
    summarise(
        in_use=sum(in_use, na.rm=TRUE),
        in_use_multiple=sum(value > 1, na.rm=TRUE)
    ) %>%
    ggplot(., aes(x=time, y=in_use)) +
        geom_bar(stat="identity") +
        geom_label(aes(label=in_use), size=3, vjust=-0.5) +
        facet_grid(. ~ date, ) +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
        scale_y_continuous(breaks=seq(0, 35, 1), limits=c(0, 35)) +
        labs(title="Number of meeting rooms in use at Oakfield House",
            y="Number of meeting rooms in use",
            x="Time")
```

In use with multiple people

```{r}
of_wt_mr %>%
    mutate(
        in_use=as.numeric(value > 0)
    ) %>%
    group_by(datetime, day, time, date) %>%
    summarise(
        in_use=sum(in_use, na.rm=TRUE),
        in_use_multiple=sum(value > 1, na.rm=TRUE)
    ) %>%
    ggplot(., aes(x=time, y=in_use_multiple)) +
        geom_bar(stat="identity") +
        geom_label(aes(label=in_use_multiple), size=3, vjust=-0.5) +
        facet_grid(. ~ date, ) +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
        scale_y_continuous(breaks=seq(0, 35, 1), limits=c(0, 35)) +
        labs(title="Number of meeting rooms in use at Oakfield House multiple occupancy",
            y="Number of meeting rooms in use by more than one person",
            x="Time")
```

How many people are in these meeting rooms?

```{r}
of_wt_mr %>% group_by(Occupancy=value) %>%
    summarise(n=n())
```

Can we pack the meetings into fewer rooms?

- Take 2025-06-17, largest number of meetings

```{r}
of_top <- subset(of_wt_mr, date == "2025-06-17")

of_top %>%
    ggplot(., aes(x=time, y=`ROOM NUMBERS`, fill=value)) +
    geom_tile() +
    scale_fill_gradient(low = "white", high = "blue")
```

```{r}
of_bot <- subset(of_wt_mr, date == "2025-06-13")

of_bot %>%
    ggplot(., aes(x=time, y=`ROOM NUMBERS`, fill=value)) +
    geom_tile() +
    scale_fill_gradient(low = "white", high = "blue")
```



## Look at door data

```{r}
library(janitor)
oh_ds <- read_xlsx(here("data/Oakfield House June 2025 Door sheet data.xlsx"), sheet="NEW FORMAT") %>%
    janitor::clean_names() %>%
    filter(!is.na(type))

oh_ds %>% group_by(number_of_people, type) %>%
    summarise(n=n(), time_spent=sum(duration_hours)) %>%
    ggplot(., aes(x=as.factor(number_of_people), y=time_spent, fill=type)) +
    geom_bar(stat="identity", position="dodge") +
    labs(title="Number of people in Oakfield House by type",
         x="Number of people",
         y="Count") +
    theme_minimal()
```


Let's look at confidential vs non-confidential meetings

```{r}
oh_ds$meetings <- NA

oh_ds$meetings[grepl(c("CTM"), oh_ds$type) | grepl(c(" CM"), oh_ds$type) | oh_ds$type == "CM"] <- "confidential"
oh_ds$meetings[oh_ds$type %in% c("NCM", "NCM, TM", "TM", "TM, FW")] <- "non-confidential"
table(oh_ds$meetings)
```



Look at teams meetings vs focused work in single occupancy

```{r}
oh_ds1 <- subset(oh_ds, number_of_people == 1)
oh_ds1 %>% group_by(type) %>%
    filter(type %in% c("TM", "FW")) %>%
    summarise(n=n(), time_spent=sum(duration_hours))
```












Start bin packing algorithm

- Remove empty rooms
- Remove full rooms
- Start with a room with 1 free slot, find a meeting with the largest occupancy that fits in the room
- Move that meeting to that room / new time
- Repeat until no more meetings can be moved
















```{r}
library(BoxPacking)

cap <- of_top$max_number
meetings <- subset(of_top, value > 0)$value
bin_pack_ffd(meetings, cap, sort = TRUE)

containers <- lapply(cap, \(i) {
    Container(length=1, height=1, width=i)
})

boxes <- lapply(meetings, \(i) {
    Box(length=1, height=1, width=i)
})

solution <- PerformBoxPacking(containers = containers,
                      boxes = boxes,
                      n_iter = 4,
                      population_size = length(boxes),
                      elitism_size = 5,
                      crossover_prob = 0.5,
                      mutation_prob = 0.5,
                      verbose = TRUE,
                      plotSolution = TRUE
                      )

```

How correlated is Planon with walkthrough data?

```{r}
oh_pl <- read_xlsx(here("data/Oakfield House Desk Bookings 9.6.25 to 20.6.25.xlsx"), sheet="All data") %>%
    janitor::clean_names() %>%
    tidyr::separate(desk_number, into=c("prefix", "room", "desk")) %>%
    mutate(date=as.Date(start_date_time, format="%d/%m/%Y")) %>%
    group_by(date, room) %>%
    summarise(
        value=n()
    )
str(oh_pl)

oh_wt_d <- oh_wt %>%
    group_by(date, `ROOM NUMBERS`, floor) %>%
    summarise(
        max_number=mean(max_number, na.rm=TRUE),
        value=max(value, na.rm=TRUE),
        occupancy_ratio=pmin(value/max_number, 1),
        office_type=ifelse(max_number == 1, "Single", "Multi")
    ) %>%
    filter(!is.na(occupancy_ratio))

oh_wt_pl <- left_join(oh_wt_d, oh_pl, by=c("date", "ROOM NUMBERS"="room")) 
oh_wt_pl$value.y[is.na(oh_wt_pl$value.y)] <- 0
oh_wt_pl %>% str

ggplot(oh_wt_pl, aes(x=value.x, y=value.y)) +
    geom_jitter(alpha=0.5) +
    geom_smooth(method="lm", se=FALSE, color="blue") +
    geom_abline(slope=1, intercept=0, linetype="dashed", color="red") +
    labs(title="Occupancy Ratio vs Planon Bookings",
         x="Walkthroughs",
         y="Planon Bookings") +
    theme_minimal()

plot(oh_wt_pl$value.x, oh_wt_pl$value.y, xlab="Walkthroughs", ylab="Planon Bookings", main="Occupancy Ratio vs Planon Bookings")
```


